Forecasting the 2024 U.S. Presidential Election: Kamala Harris' Projected Victory

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd

### Predictive Variables
In this section, we outline the two key predictor variables used in our analysis: 'state', which captures the geographical context of the polling data, and 'end_date', which reflects the timing of the polls relative to the election, allowing us to assess trends in voter support over ti
#### State
The state ('state') variable indicates the U.S. state where the poll was conducted or targeted. In the context of the U.S. presidential election, the outcome is frequently determined by several key battlegroud states. In the current 2024 election cycle, there are seven battleground states that are pivotal in the contest between Kamala Harris, representing the Democratic Party, and Donald Trump, representing the Republican Party. To enhance our model and refine our predictions, we focused exclusively on these seven battleground states. The polling results from each of these states play a crucial role in predicting electoral outcomes, as the percentage of support is evaluated on a state-by-state basis.
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-state-distribution
#| fig-cap: "Bar plot showing the distribution of observations across key battleground states in the 2024 United States Presidential Election polling data. The plot illustrates the count of polling observations for each state, highlighting states with more frequent polling activity. This visualization aids in understanding the level of engagement and data collection in each battleground state leading up to the election."

@@

### Other Variables
This section details additional variables that provide essential context for our analysis, including numeric grade, which evaluates the quality of the polls; transparency score, indicating the openness of poll methodologies; pollster, identifying the organizations conducting the polls; and start date, which denotes the initiation of the polling period for each survey.
#### Numeric Grade
Numeric Grade ('numeric_grade') represents the quality of each poll, as rated by [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy. In the context of our paper, numeric grade allows us to weigh the reliability of different polls. By including numeric grade as a variable in our model, we aim to account for variations in poll quality, ensuring that polls with higher ratings (indicating higher reliability) contribute more significantly to our predictions of voter support. @fig-numgrade-distribution displays the distribution of numeric grades for the polls included in our analysis, ranging from 2.5 to 3.0. Polls with higher grades are generally seen as more methodologically rigorous and transparent, which helps improve the reliability of our model’s predictions. The distribution shows several peaks, indicating that certain grade levels (around 2.7, 2.8, and 3.0) are more common among the polling data we collected.
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-numgrade-distribution
#| fig-cap: "Histogram illustrating the distribution of polling end dates for key battleground states in the lead-up to the 2024 United States Presidential Election. Displays the frequency of polls conducted over time, providing insight into the timing of polling activities and potential trends in voter sentiment as Election Day approaches. This visualization is crucial for understanding how polling data correlates with the election timeline."

@@

## Model Summary
### Harris Model Summary
```{r}
#| echo: false
#| eval: true
#| label: tbl-harris-fixed-effects
#| tbl-cap: "Fixed Effects Summary for Harris Model."
#| warning: false

@@

The model summary for Kamala Harris indicates that the intercept estimate is $-0.1508$ with a standard error of $0.0279$. This intercept value suggests a negative baseline log-odds for her victory, reflecting the starting point of her support in the absence of any other predictors. The model shows that the coefficients for the natural spline terms ($nsendate_{num}$) are particularly informative, with the first spline term ($nsendate_{num\_df41}$) yielding a positive estimate of $0.1225$, suggesting that as time progresses, Harris’s probability of winning increases, particularly during key moments captured by the polling data. This is also reflected in the confidence intervals, with the upper limit of the confidence interval for this term reaching $0.1903$. In contrast, the estimates for the second spline term ($nsendate_{num\_df42}$) indicate a smaller effect, suggesting fluctuations in her support that may warrant further investigation.

### Trump Model Summary
```{r}
#| echo: false
#| eval: true
#| label: tbl-trump-fixed-effects
#| tbl-cap: "Fixed Effects Summary for Trump Model."
#| warning: false

# Display the Trump model summary table
knitr::kable(trump_summary) %>%
  kable_styling(full_width = FALSE, position = "center")
```
In comparison, the model summary for Donald Trump presents a different picture. The intercept for Trump's model is $-0.0923$ with a standard error of $0.0306$, indicating a similarly negative baseline log-odds of victory as Harris. However, Trump's model demonstrates more pronounced increases in support as reflected by the estimates for the natural spline terms. For example, the second spline term ($nsendate_{num\_df42}$) has a positive coefficient of $0.0803$, highlighting a significant upward trend in support over time. This trend is further supported by the confidence interval, which ranges from $0.0179$ to $0.1426$, indicating a substantial increase in his predicted probabilities as the election approaches. The fourth spline term ($nsendate_{num\_df44}$) also shows a positive coefficient of $0.0789$, suggesting that the support for Trump remains resilient over the polling period. Overall, both models reveal complex interactions between time and voter sentiment, with implications for the electoral dynamics leading up to the election.

>>>>

### Predictive Variables
In this section, we outline the two key predictor variables used in our analysis: 'state', which captures the geographical context of the polling data, and 'end_date', which reflects the timing of the polls relative to the election, allowing us to assess trends in voter support over time.

#### State
The state ('state') variable indicates the U.S. state where the poll was conducted or targeted. In the context of the U.S. presidential election, the outcome is frequently determined by several key battlegroud states. In the current 2024 election cycle, there are seven battleground states that are pivotal in the contest between Kamala Harris, representing the Democratic Party, and Donald Trump, representing the Republican Party. To enhance our model and refine our predictions, we focused exclusively on these seven battleground states. The polling results from each of these states play a crucial role in predicting electoral outcomes, as the percentage of support is evaluated on a state-by-state basis.
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-state-distribution
#| fig-cap: "Bar plot showing the distribution of observations across key battleground states in the 2024 United States Presidential Election polling data. The plot illustrates the count of polling observations for each state, highlighting states with more frequent polling activity. This visualization aids in understanding the level of engagement and data collection in each battleground state leading up to the election."

@@

### Other Variables
This section details additional variables that provide essential context for our analysis, including numeric grade, which evaluates the quality of the polls; transparency score, indicating the openness of poll methodologies; pollster, identifying the organizations conducting the polls; and start date, which denotes the initiation of the polling period for each survey.

#### Numeric Grade
Numeric Grade ('numeric_grade') represents the quality of each poll, as rated by [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy. In the context of our paper, numeric grade allows us to weigh the reliability of different polls. By including numeric grade as a variable in our model, we aim to account for variations in poll quality, ensuring that polls with higher ratings (indicating higher reliability) contribute more significantly to our predictions of voter support. @fig-numgrade-distribution displays the distribution of numeric grades for the polls included in our analysis, ranging from 2.5 to 3.0. Polls with higher grades are generally seen as more methodologically rigorous and transparent, which helps improve the reliability of our model’s predictions. The distribution shows several peaks, indicating that certain grade levels (around 2.7, 2.8, and 3.0) are more common among the polling data we collected.
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-numgrade-distribution
#| fig-cap: "Histogram illustrating the distribution of polling end dates for key battleground states in the lead-up to the 2024 United States Presidential Election. Displays the frequency of polls conducted over time, providing insight into the timing of polling activities and potential trends in voter sentiment as Election Day approaches. This visualization is crucial for understanding how polling data correlates with the election timeline."

@@

## Model Summary
### Trump Model Summary
```{r}
#| echo: false
#| eval: true
#| label: tbl-trump-fixed-effects
#| tbl-cap: "Fixed Effects Summary for Trump Model."
#| warning: false

# Display the Trump model summary table
knitr::kable(trump_summary) %>%
  kable_styling(full_width = FALSE, position = "center")
```

### Harris Model Summary
```{r}
#| echo: false
#| eval: true
#| label: tbl-harris-fixed-effects
#| tbl-cap: "Fixed Effects Summary for Harris Model."
#| warning: false

@@
